### 2015-07-02 narrative/enabling_tech.tex
minor changes.
move some of the Cepth and ADIOS part into the enabling technologies, and try to re-organize to show the focus of this proposal is on theme 2.

\section*{Enabling Technologies}

Significant existing work will serve as the foundation upon which this project
will be performed.

@@

>>>>

\section{Enabling Technologies}
\label{sec:enabling-technologies}
Significant existing work will serve as the foundation upon which this project
will be performed.

@@

In this project we will extend ADIOS~\cite{adios} with new APIs and novel techniques
to coordinate with the storage back-end. The key idea is offering the ability to
negotiate with the storage system to agree on the tradeoff between data quality
and response time. We will test our ideas on the next generation OLCF,
SNL/LANL, and NERSC systems.
%
We plan to leverage the many inovations we discovered in 
the Lightweight File Systems~\cite{lwfs} (LWFS) project and the Ceph storage system project (ceph.com).
In the LWFS project, Sandia sought to
demonstrate stripping down a file system to the bare minimum features required
for any storage system. Then, additional functionality and semantics, such as
POSIX-style directories and consistency requirements, were added in using
auxiliary services separate from the stack or by layering providing a new
interface to the storage system. The project's first phase demonstrated that
using a core object storage system with integrated security providing
authentication and authorization services, customized storage interfaces could
be developed with exactly the overhead required for different applications. The
second phase of LWFS, known as Sirocco~\cite{sirocco}, is investigating how to
manage storage devices for efficient and scalable storage performance.  Sirocco
has been in process for three years and has developed a usable storage infrastructure.

The Ceph project (ceph.com) started at UC Santa Cruz in 2004 and
was initially funded by DOE/NNSA involving LLNL, LANL, and Sandia,
with the goal to create an object-based parallel file system that
addresses the well-known metadata service bottleneck of creating
and accessing 1000s of files comprising a checkpoint~\cite{weil:osdi06}.
However, after the Ceph kernel client was pulled into the Linux
mainline in 2010, the focus of Ceph shifted to its object-based
storage subsystem, also called Reliable Autonomous Distributed
Object Store (RADOS), which now is the dominant storage system in
OpenStack production-level deployments and was recently selected
as the new backend for Flickr and Yahoo! Cloud Services. In this
project we will leverage two of the key innovations of Ceph: (1)
Controlled Replication Under Scalable Hashing (CRUSH)~\cite{weil:sc06}
will help making resource identification and predictable performance
scalable, and (2) extensible storage object classes~\cite{watkins:ucsctr15}
will provide the main mechanisms of introducing new APIs,
metadata vocabulary, and quality of service into the storage stack.

### 2015-07-02 narrative/intro.tex
move some of the Cepth and ADIOS part into the enabling technologies, and try to re-organize to show the focus of this proposal is on theme 2.

Our approach targets the expected characteristics of exascale storage
hardware. The storage layer will be partitioned into multiple heterogenous
tiers with vastly different performance characteristics. The layers
will be further differentiated by the constraints on capacity
and data lifetime within each layer. Tape archival storage will still maintain
data long term, but access to this data will be orders of magnitude slower
than the next layer and naively accessing data from archival storage will
greatly impact productivity.
%
{\em The purpose of computing is insight, not numbers}~\cite{Hamming:book}.
Richard Hamming's remark, made over 50 years ago at the dawning of the age of
large scale scientific computation, is even more relevant today as we prepare
for scientific simulation at the exascale. Hamming was a pioneer in the fields
of coding theory, scientific computing and information theory who would
instantly sympathize with the central objective of this proposal--minimizing
the time to knowledge.  Our approach in this proposal is to enable researchers
to place more knowledge into the SSIO system so that data can be recast from a bunch of bytes
to information which the system can then place and optimize the time to gain knowledge from exascale simulations.
We first motivate this by a problem our team has been working on in the past six months, and then recast this problem into
a series of research questions and challenges which must be addressed in order to optimize the SSIO system for exascale science.

The traditional approach to transforming results of scientific computation into
knowledge consists of storing the numerical output of a simulation on disk,
followed by a post-processing or analysis stage that might take place over
several ensuing weeks or months. The overarching objective is to gain insight,
or knowledge.  As already mentioned, the anticipated scale of computations
means that this workflow is simply no longer viable.
%
The overarching idea is that data from applications can be characterized by different
levels of importance and different level of importance over time. 
Codes such as the XGC1 simulation from C. S. Chang (PPPL), which is one of largest users of Leadership Class Facilities (over 300M hours at ANL, NERSC, and ORNL) has launched a series of simulations which need to write out 100 PB of data on the Titan system. Due to resource limitations, the simulation will only write about 10 PB, but
when our team considered not only the time it would take to write out 10 PB in 10 days of run, but also where we would store this, which had to be tape since it would take
over half of the file system, we decided that we couldn't write this amount of data. In other words we had to make decisions of which data to save from the original 100 PB.
The first thing that we had to discover is how much space we could have on the parallel file system at any given time, and how much time it would take to archive 1 PB of data,
which turned out to be 10 days, and we could only store 2 PB of data at once. This quickly led us to 1) decide which pieces of data to save from the large amount of data; in 
other words, we had to prioritize which were the most important pieces. We had to further breakdown the large data into two categories, data which we think we will touch after
the simulation is complete ( after the 10 days of wallclock time was finished), and what we would do with this data (the access pattern of the large data). We then had to 
discover application specific data reduction techniques, and encode that into our solution, on top of ADIOS. We then had to implement a crude discovery system to understand
which pieces of data were stored on the parallel file system and which pieces moved over to PPPL, and which pieces were on tape.  This solution worked well for one day
but after the second day more problems occurred, the variability of the time it took to write the data made the I/O prohibitively costly, because even though we had the
majority of the OLCF, the file system was shared and sometimes the writing time took less 100 seconds, and at other times it was over 1,000 seconds. We had no method
to monitor this time and then write less data when the time was going to be longer.   After the simulation is complete we also need to visualize this reduced data, and 
an additional problem appeared; just to get a first look at the visualization data seemed like it could take over a month because some of the timesteps were on tape and other
timesteps were on the file system. We had no method to just look at the data quickly and understand which regions of the simulation to focus on.

@@

In this project we will extend ADIOS~\cite{adios} with new APIs and novel techniques
to coordinate with the storage back-end. The key idea is offering the ability to
negotiate with the storage system to agree on the tradeoff between data quality
and response time. We will test our ideas on the next generation OLCF,
SNL/LANL, and NERSC systems.
%
We plan to leverage the many inovations we discovered in 
the Lightweight File Systems~\cite{lwfs} (LWFS) project and the Ceph storage system project (ceph.com).
In the LWFS project, Sandia sought to
demonstrate stripping down a file system to the bare minimum features required
for any storage system. Then, additional functionality and semantics, such as
POSIX-style directories and consistency requirements, were added in using
auxiliary services separate from the stack or by layering providing a new
interface to the storage system. The project's first phase demonstrated that
using a core object storage system with integrated security providing
authentication and authorization services, customized storage interfaces could
be developed with exactly the overhead required for different applications. The
second phase of LWFS, known as Sirocco~\cite{sirocco}, is investigating how to
manage storage devices for efficient and scalable storage performance.  Sirocco
has been in process for three years and has developed a usable storage infrastructure.

The Ceph project (ceph.com) started at UC Santa Cruz in 2004 and
was initially funded by DOE/NNSA involving LLNL, LANL, and Sandia,
with the goal to create an object-based parallel file system that
addresses the well-known metadata service bottleneck of creating
and accessing 1000s of files comprising a checkpoint~\cite{weil:osdi06}.
However, after the Ceph kernel client was pulled into the Linux
mainline in 2010, the focus of Ceph shifted to its object-based
storage subsystem, also called Reliable Autonomous Distributed
Object Store (RADOS), which now is the dominant storage system in
OpenStack production-level deployments and was recently selected
as the new backend for Flickr and Yahoo! Cloud Services. In this
project we will leverage two of the key innovations of Ceph: (1)
Controlled Replication Under Scalable Hashing (CRUSH)~\cite{weil:sc06}
will help making resource identification and predictable performance
scalable, and (2) extensible storage object classes~\cite{watkins:ucsctr15}
will provide the main mechanisms of introducing new APIs,
metadata vocabulary, and quality of service into the storage stack.

>>>>

Our approach targets the expected characteristics of exascale storage
hardware. The storage layer will be partitioned into multiple heterogeneous
tiers with vastly different performance characteristics. The layers
will be further differentiated by the constraints on capacity
and data lifetime within each layer. Tape archival storage will still maintain
data long term, but access to this data will be orders of magnitude slower
than the next layer and naively accessing data from archival storage will
greatly impact productivity.
%
{\em The purpose of computing is insight, not numbers}~\cite{Hamming:book}.
Richard Hamming's remark, made over 50 years ago at the dawning of the age of
large scale scientific computation, is even more relevant today as we prepare
for scientific simulation at the exascale. Hamming was a pioneer in the fields
of coding theory, scientific computing and information theory who would
instantly sympathize with the central objective of this proposal--minimizing
the time to knowledge.  Our approach in this proposal is to enable researchers
to place more knowledge into the SSIO system so that data can be recast from a bunch of bytes
to information which the system can then place and optimize the time to gain knowledge from exascale simulations.
We first motivate this by a problem our team has been working on in the past six months, and then recast this problem into
a series of research questions and challenges which must be addressed in order to optimize the SSIO system for exascale science.

The traditional approach to transforming results of scientific computation into
knowledge consists of storing the numerical output of a simulation on disk,
followed by a post-processing or analysis stage that might take place over
several ensuing weeks or months. The overarching objective is to gain insight,
or knowledge.  As already mentioned, the anticipated scale of computations
means that this workflow is simply no longer viable.
%
\section{Scalable Storage Software Infrastructure}
Our focus in this proposal is in theme 2: Scalable Storage Software Infrastructure. We will focus on a state of the art
SSIO solution to solve problems which researchers are beginning to face and will make science on exascale systems
problematic without new insights and solutions.

The overarching idea is that data from applications can be characterized by different
levels of importance and different level of importance over time. 
Codes such as the XGC1 simulation from C. S. Chang (PPPL), which is one of largest users of Leadership Class Facilities (over 300M hours at ANL, NERSC, and ORNL) has launched a series of simulations which need to write out 100 PB of data on the Titan system. Due to resource limitations, the simulation will only write about 10 PB, but
when our team considered not only the time it would take to write out 10 PB in 10 days of run, but also where we would store this, which had to be tape since it would take
over half of the file system, we decided that we couldn't write this amount of data. In other words we had to make decisions of which data to save from the original 100 PB.
The first thing that we had to discover is how much space we could have on the parallel file system at any given time, and how much time it would take to archive 1 PB of data,
which turned out to be 10 days, and we could only store 2 PB of data at once. This quickly led us to 1) decide which pieces of data to save from the large amount of data; in 
other words, we had to prioritize which were the most important pieces. We had to further breakdown the large data into two categories, data which we think we will touch after
the simulation is complete ( after the 10 days of wallclock time was finished), and what we would do with this data (the access pattern of the large data). We then had to 
discover application specific data reduction techniques, and encode that into our solution, on top of ADIOS. We then had to implement a crude discovery system to understand
which pieces of data were stored on the parallel file system and which pieces moved over to PPPL, and which pieces were on tape.  This solution worked well for one day
but after the second day more problems occurred, the variability of the time it took to write the data made the I/O prohibitively costly, because even though we had the
majority of the OLCF, the file system was shared and sometimes the writing time took less 100 seconds, and at other times it was over 1,000 seconds. We had no method
to monitor this time and then write less data when the time was going to be longer.   After the simulation is complete we also need to visualize this reduced data, and 
an additional problem appeared; just to get a first look at the visualization data seemed like it could take over a month because some of the timesteps were on tape and other
timesteps were on the file system. We had no method to just look at the data quickly and understand which regions of the simulation to focus on.

@@
